Cyclist video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda

### 2024-05-03 README.qmd
Update readme

# In preparation for reading-in all data:
date_str = format(Sys.Date(), "%Y-%m")
file_name = paste0("data/west-yorkshire/operation_snap_", date_str, ".csv")
```

@@

```{r}
#| eval: false
#| echo: true
u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
d = opsnap:::download_and_read(u)
names(d)
# Old names:
#  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
#  [3] "OFFENDER VEHICLE MODEL"  "OFFENDER VEHICLE COLOUR"
#  [5] "OFFENCE"                 "DISTRICT"               
#  [7] "DISPOSAL"                "DATE OF SUBMISSION"     
#  [9] "...9"                    "OFF LOCATION"
# New names:
# [1] "mode"     "make"     "model"    "colour"   "offence"  "district" "disposal"
# [8] "date"     "location"
```

@@

```{r}
#| echo: false
d_all = read_csv(file_name)
d_all = d_all |>
  mutate(
    mode = tolower(mode),
    offence = tolower(offence),
    disposal = tolower(disposal)
  )
d_all |>
  head(3) |>
  knitr::kable()
```

@@

```{r}
d_all_monthly = d_all |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "all")
d_offence = d_all |>
  opsnap:::filter_offence_nas()
d_offence_monthly = d_offence |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete_offence")
d = d_offence |>
  opsnap:::filter_location_nas()
d_complete_monthly = d |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete")
d_monthly = bind_rows(d_all_monthly, d_offence_monthly, d_complete_monthly)
d_monthly |>
  ggplot() +
  geom_line(aes(month, n, colour = records), alpha = 0.5, size = 2) +
  # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
  labs(title = "Number of records in West Yorkshire Police Operation Snap data",
       x = "Date",
       y = "Number of records per month") +
  theme_minimal()
```

As shown in the graph above, `r round(nrow(d_offence) / nrow(d_all) * 100, 1)`% have values for the 'offence' column. 
Many records lack either an offence or a location, leaving only `r round(nrow(d) / nrow(d_all) * 100, 1)`% or `r nrow(d)` complete records.

The breakdown of records by mode of transport (of the observer) is shown below:

```{r}
d_all |>
  count(mode, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  knitr::kable()
```

@@

```{r}
d_all |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

The equivalent table excluding records with missing offence data is shown below:

```{r}
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_offence)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

The equivalent for cyclists, with the least common offences categorised as 'other', is shown below:

```{r}
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  filter(mode == "cyclist") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_offence)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  mutate(
    offence = ifelse(n < 20, "other", offence)
  ) |>
  group_by(offence) |>
  summarise(n = sum(n), n_hybrid = n()) |>
  arrange(n_hybrid, desc(n)) |>
  select(-n_hybrid) |>
  mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
  knitr::kable()
```

@@

```{r}
d_all |>
  count(disposal, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  knitr::kable()
```

There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:

```{r}
#| echo: false
d |>
  count(location, sort = TRUE) |>
  mutate(percent_records = n / nrow(d)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

@@

```{r}
#| eval: false
#| echo: true
d_sample = d[1:5, ]
d_sf = opsnap:::op_geocode(d_sample)
mapview::mapview(d_sf)
```

```{r}
#| eval: false
#| echo: false
# Geocode all data:
d_geocoded = opsnap:::op_geocode(d)
table(d_geocoded$address) |>
  sort() |>
  tail(1)
# d_geometries = d_geocoded |>
#   filter(address != "NA, West Yorkshire") |>
  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
d_sf = sf::st_sf(
  d,
  geometry = d_geometries$geometry
)
west_yorkshire = pct::get_pct_zones("west-yorkshire")
sf::sf_use_s2(FALSE)
d_sf_wy = d_sf[west_yorkshire, ]
proportion_outside_wy = 1 - nrow(d_sf_wy) / nrow(d_sf)
# Sanity check
d_sf_wy |>
  sample_n(1000) |>
  # mapview::mapview()
  plot()
# Save to file:
sf::write_sf(d_sf_wy, paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
```

@@

```{r}
d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
d_projected = sf::st_transform(d_sf, 27700)
library(terra)
raster_template = rast(ext(d_projected), res = 500)
raster_count = rasterize(d_projected, raster_template, fun = "length")
plot(raster_count)
```

@@

```{r}
#| eval: false
#| echo: false
table(d$offence) |>
  sort()
```

@@

```{r}
#| eval: false
#| echo: false
d |>
  # Reduce nchar of offence
  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
  group_by(offence) |>
  # Count number of rows in each group
  mutate(n = n()) |>
  filter(n > nrow(d) / 50)|>
  ggplot() +
  geom_bar(aes(offence)) +
  # Make x labels vertical
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

>>>>

# In preparation for reading-in all data:
# date_str = format(Sys.Date(), "%Y-%m")
date_str = format(as.Date("2024-04-01"), "%Y-%m")
file_name = paste0("data/west-yorkshire/operation_snap_", date_str, ".csv")
```

@@

```{r}
#| eval: false
#| echo: true
u = "https://www.westyorkshire.police.uk/sites/default/files/2024-01/operation_snap_oct-dec_2023_0.xlsx"
d = opsnap:::download_and_read(u)
names(d_with_location)
# Old names:
#  [1] "REPORTER TRANSPORT MODE" "OFFENDER VEHICLE MAKE"  
#  [3] "OFFENDER VEHICLE MODEL"  "OFFENDER VEHICLE COLOUR"
#  [5] "OFFENCE"                 "DISTRICT"               
#  [7] "DISPOSAL"                "DATE OF SUBMISSION"     
#  [9] "...9"                    "OFF LOCATION"
# New names:
# [1] "mode"     "make"     "model"    "colour"   "offence"  "district" "disposal"
# [8] "date"     "location"
```

@@

```{r}
#| echo: false
d_all = read_csv(file_name)
d_all = d_all |>
  mutate_if(is.character, tolower)
d_all |>
  head(3) |>
  knitr::kable()
```

@@

```{r}
d_all_monthly = d_all |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "All")
d_with_offence = d_all |>
  opsnap:::filter_offence_nas()
# nrow(d_all) - nrow(d_with_offence) # only 1 record -> ignore
d_with_location = d_all |>
  opsnap:::filter_location_nas()
d_with_location_monthly = d_with_location |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "With location")
d_monthly = bind_rows(d_all_monthly, d_with_location_monthly)
d_monthly |>
  ggplot() +
  geom_line(aes(month, n, colour = records), alpha = 0.5, size = 2) +
  # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
  labs(title = "Number of records in West Yorkshire Police Operation Snap data",
       x = "Date",
       y = "Number of records per month") +
  theme_minimal() +
  ylim(c(0, NA))
```

As shown in the graph above not all (`r round(nrow(d_with_location) / nrow(d_all) * 100, 1)`%) records have values for the 'location' column.

The breakdown of all records by mode of transport (of the observer) is shown below:

```{r}
d_all |>
  count(mode, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
  arrange(desc(n)) |>
  knitr::kable()
```

@@

```{r}
d_all |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

The equivalent table excluding records with missing location data is shown below:

```{r}
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_with_location)) |>
  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

The equivalent for cyclists, with location present and the least common offences categorised as 'other', is shown below:

```{r}
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  filter(mode == "cyclist") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_with_location)) |>
  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
  mutate(
    offence = ifelse(n < 20, "other", offence)
  ) |>
  group_by(offence) |>
  summarise(n = sum(n), n_hybrid = n()) |>
  arrange(n_hybrid, desc(n)) |>
  select(-n_hybrid) |>
  mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
  knitr::kable()
```

@@

```{r}
d_all |>
  count(disposal, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
  arrange(desc(n)) |>
  knitr::kable()
```

There are `r unique(d_with_location$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:

```{r}
#| echo: false
d_with_location |>
  mutate(location = tolower(location)) |> 
  count(location, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_with_location)) |>
  mutate(percent_records = scales::percent(percent_records, accuracy = 0.1)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

@@

```{r}
#| eval: false
#| echo: true
d_sample = d_with_location[1:5, ]
d_sf = opsnap:::op_geocode(d_sample)
mapview::mapview(d_sf)
```

```{r}
#| eval: false
#| echo: false
# Geocode all data:
d_geocoded = opsnap:::op_geocode(d_with_location)
table(d_geocoded$address) |>
  sort() |>
  tail(1)
# d_geometries = d_geocoded |>
#   filter(address != "NA, West Yorkshire") |>
  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
d_sf = sf::st_sf(
  d_with_location,
  geometry = d_geometries$geometry
)
west_yorkshire = pct::get_pct_zones("west-yorkshire")
sf::sf_use_s2(FALSE)
d_sf_wy = d_sf[west_yorkshire, ]
proportion_outside_wy = 1 - nrow(d_sf_wy) / nrow(d_sf)
# Sanity check
d_sf_wy |>
  sample_n(1000) |>
  # mapview::mapview()
  plot()
# Save to file:
sf::write_sf(d_sf_wy, paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
```

@@

```{r}
d_sf = sf::read_sf(paste0("data/west-yorkshire/operation_snap_geocoded_", date_str, ".gpkg"))
```

```{r}
#| eval: false
d_sf_aggregated = d_sf |> 
  group_by(location) |> 
  count(mode)
library(tmap)
tmap_mode("view")
d_sf |> 
  sample_n(1000) |> 
  tm_shape() +
  tm_dots()
```

```{r}
d_projected = sf::st_transform(d_sf, 27700)
library(terra)
raster_template = rast(ext(d_projected), res = 500)
raster_count = rasterize(d_projected, raster_template, fun = "length")
# mapview::mapview(raster_count)
plot(raster_count)
```

@@

```{r}
#| eval: false
#| echo: false
table(d_with_location$offence) |>
  sort()
```

@@

```{r}
#| eval: false
#| echo: false
d |>
  # Reduce nchar of offence
  mutate(offence = stringr::str_sub(offence, 1, 60)) |>
  group_by(offence) |>
  # Count number of rows in each group
  mutate(n = n()) |>
  filter(n > nrow(d_with_location) / 50)|>
  ggplot() +
  geom_bar(aes(offence)) +
  # Make x labels vertical
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
